Untwisting the Boerdijk-Coxeter Helix

### 2017-05-26 doc/TetrahelixUntwist.tex
getting closer to final

\begin{abstract}
  The Boerdijk-Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra forming a helical column.  Considering the edges of
  these tetrahedra as structural members, the resulting structure is attractive and
  inherently rigid, and therefore interesting to architects,
  mechanical engineers, and roboticists.  A formula is developed that matches the
  visually apparent helices forming the outer rails of the BC helix.
  This formula is generalized to a formula convenient to designers.
  Formulae for 
  computing the
  parameters that give edge-length minimax-optimal tetrahelices
  are given, defining a continuum of tetrahelices of varying curvature.
  The endpoints of the optimality of this continuum are the BC helix and
  a structure of zero curvature, the \emph{equitetrabeam}.
  Numerically finding the rail angle from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed. 
  An interactive tool for such design and experimentation is provided: \url{https://pubinv.github.io/tetrahelix/}.
  A formula for the inradius of optimal tetrahelices is given.
  Utility for static and variable geometry
  truss/space frame design and robotics is discussed.
\end{abstract}

@@

Imagining \cref{fig:closeup,fig:closeuporthog} as a static mechanical structure,
we observe that it is useful to the mechanical engineer or
roboticist because the structure remains an inherently rigid,
omni-triangulated space frame, which is
mechanically strong.
Then we can imagine that each static edge is replaced with an
actuator that can dynamically become shorter or longer in response to electronic control,
and the vertices are joints that support sufficient angular displacement
for this to be possible. An example of such a machine is a Tetrobot,
shown in \cref{fig:tetrobot}.

A BC helix does not rest stably on a plane. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
developed here is available in JavaScript and demonstrated by an interactive
design website \url{https://pubinv.github.io/tetrahelix/}\cite{readtetrahelix},
from which \cref{fig:closeup,fig:closeuporthog,fig:series}
are taken.

@@

The apparent rotation of a vertex on an outer-edge, that is $\vec{V}(n)$ relative from $\vec{V}(n+3)$
for any integer $n$
in \eqref{graycoxeter}, is $3 \theta_{bc} - 2\pi$).

This formula defines a helix, but it is not any of the apparent helices, or \emph{rail} helices, of the
BC helix, but rather one that winds three times as rapidly through all
nodes. To a designer of tetrahelices, it is more natural to think of
the three helices which are visually apparent, that is, those three
which are closely approximated by the outer edges or rails of
the BC helix. We think of each of these three rails as being a different color: red, blue, or yellow.
This situation is illustrated in \cref{fig:helixcomparison}, wherein the black helix represents that
generated by \cref{graycoxeter},
and the colored helices are generated by \cref{eq:colored}.

@@

Such a helix can be written:
\begin{equation}
  \label{eq:colored}
 \vec{H}_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})}  $\\
   $ r_{bc} \sin{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})} $\\
   $ 3 h_{bc} (n + c/3)  $
  \end{tabular}
  \right ],
\text{where}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta_{bc} = \arccos(-2/3) $ \\
  \end{tabular}      
\end{equation}

@@

The $\vec{H}_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho_{bc}$ rather than $\theta_{bc}$.
Intuitively we seek an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$.
We expand 
the expressions $\theta_{bc}$ and $\rho_{bc}$ in \eqref{eq:colored} and seek to isolate the term $c 2\pi/3 $.
\begin{align*}
 c \theta_{bc}  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (c/3)  (3 \theta_{bc})  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
 (c/ 3) ((3 \theta_{bc} - 2 \pi)  + 2 \pi) &= \text{\{definition of $\rho_{bc}$\ ...\}} \\  
  (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{\{algebra...\}} \\  
c  ( \rho_{bc} +  2 \pi) /3  \\
\end{align*}

@@

The pitch is less than the number of tetrahedra because the tetrahedra
edges are not parallel to the axis of the tetrahelix.  It is a famous and interesting result
that the pitch is irrational. A BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of developing a rational period by arranging solid tetrahedra by relaxing the face-to-face matching
has been explored\cite{sadler2013periodic}. 
We develop below slightly irregular edge lengths that support, for example, a pitch of precisely 12
tetrahedra in one revolution which would allow an architect to
design a column having the top and bottom tetrahedra in the same relationship to the capital and the basis to the
viewer.

@@

In the proof below we find useful to consider projection diagrams that are the axial projection
of a tetrahelix onto the $XY$-plane. \Cref{fig:projectiondiagram} is an example of such a diagram.
\begin{lemma}
  If the rail angle $0 < \rho < \pi$ is a rational multiple of $\pi$, then the projection of
  edges long a helix of that rail angle  along the $z$-axis onto the $XY$-plane form a
  regular polygon of 3 or more sides, else they fill in a complete circle.
  \label{lemma:ngon}
\end{lemma}
\begin{proof}
  All points lying on a helix projected along the axis lie on a circle in the $XY$-plane.
  Helices are periodic in the $z$ dimension modulo $2\pi$.
  If $2\pi/\rho$ is irrational, the projection onto the $XY$-plane will contain
  an unbounded number of points on a circle.
  If and only if $2\pi/\rho$ is rational, the projection onto
  the $XY$-plane will contain a finite number of points.
  Because $\pi$ is transcendental and irrational,
  $2\pi/\rho$ is rational if and only $\rho = a\pi/b$, where $a$ and $b$ are integers
  and without loss of generality $a$ and $b$ are coprime.
  Since $\rho < \pi$, therefore $a < b$. Also, $\rho > 0$, therefore $a > 0$.
  The number of points in the projection is $2b$ if $a$ is odd, and $b$ if $a$ is even.
  This polygon has at least 3 sides, since either $\rho$ is irrational or $b > a$, and therefore
  $b >= 2$. If $a/b = 1/2$, the projection is a square, which has four sides. 
\end{proof}

\begin{theorem}
  Any optimal tetrahelix with a rail angle of magnitude less than $\pi$ has all three
  axes conincident.
  \label{thm:conincident}
\end{theorem}
\begin{proof}
  Case 1: Suppose that $\rho$ is zero.
  Each helix has zero curvature, that is, is a straight line. These lines are eqivalent
  to some three degenerate helices, possibly with different radii, so long as there is a phase
  term in the defintion of the helix, as in \eqref{eq:colored}. We later show the
  radii must be eqivalent.

  Case 2: Suppose that $\rho$ is positive but less than $\pi$.
  In this case each rail helix has
  curavature. The projection of points in the $XY$ plane creates a figure
  guaranteed to have points on either side of any line through the axis of such
  a helix, because the figure is either an $n$-gon or a circle \cref{lemma:ngon}.
  We show that the three helices share a common axis.

  Without loss of generality define the Red helix to have its axis on the $z$-axis.
  Since either a Red-to-Yellow or a Red-to-Blue edge is either a minimum or a maximum,
  without loss of generatlity define the Blue helix to be a helix that
  has an edge connection to the Red helix that is either a maximum or a minimum.
  Let $B'$ be a translation in the $XY$-plane of the blue helix $B$  so that its axis is the $z$-axis and
  conincident with the red helix $R$. Let $D$ be the distance between the axis of the Blue helix
  $B$ and $B'$. We will show that if $D>0$ then $B$ ``wobbles'' in a way that cannot be optimal.
  Define a wobble vector by:
  \begin{equation*}
    \vec{W}(n) = \vec{B}(n) - \vec{B}'(n) \text{ .}
  \end{equation*}
  where $\vec{B}(n)$ and $\vec{B'}(n)$ is the cartesian vector
  $\begin{bmatrix}
    x \\
    y 
  \end{bmatrix}$
  for the projection of the $n$th vertex of $B$ and $B'$. 
  Note that $|| \vec{R}(n) - \vec{B'}(n+k)||$ (the Euclidean distance of the vertices)
  is a constant for any $k$, because $R$ and $B'$ have the
  same pitch and the same axis, even if they do not have the same radius.

@@

    This process can be carried out on both the Blue and Yellow helices
    (perhaps simultaneously) until $\vec{W}(n)$ is
    zero for both, finding a tetrahelix of improved overall minimax value at each step.
    So a tetrahelix is optimal only when $\vec{W}(n) = 0$, and therefore when $D=0$
    $\vec{B}(n) = \vec{B'}(n)$, and all three axes are conincident.
\end{proof}

@@

In principle in any three helices generated with \cref{eq:triple}
has at most nine distinct edge length classes. Each edge that connects
two rails potentially has a longer length and shorter length we donote with
a $+$ or $-$. So the classes are $\{ RR, BB, YY, RB_+, RB_-, BY_+, BY_-, RY_+, RY_- \}$.
If when projecting all vertices onto the $z$-axis (dropping the $x$ and $y$ coordinates), the interval
defined by the $z$ axis value of its endpoints contains no other vertices,
we call it a \emph{one-hop} edge, and if it does contain another vertex we
call it a \emph{two-hop} edge, as illustrated in \cref{fig:naming}.
Then there are 
3 rail edges $\{ RR, BB, YY\}$,
3 one-hop lengths $\{ RB_-, BY_-, RY_- \}$ between each pair of 3 rails,
and 3 two-hop
lengths $\{ RB_{+}, BY_+, RY_+ \}$ between each pair of 3 rails,
where the two-hop length is at least the one-hop length.
However, if we generate the three helices symmetrically with \cref{eq:triple},
many of these lengths will be the same.
In fact, it is possible that there
will be only two distinct such classes, or even one, in the purely regular BC helix.

\begin{theorem}
  Optimal tetrahelices have the same radii for all three helices.
\end{theorem}
\begin{proof}
To prove this we exhibit a symmetric tetrahelix (not yet shown to be optimal) which
happens to be a triple helix, that has
the property that all rail edges are equal to all one-hop edges and all two-hop
edges are equal to each other. 
Observe that although we have not get given the formula for the
radii of such a triple helix, we observe there are some values for $r$ and $\alpha$, and $\phi$
in \cref{eq:triple}
for which all the three helices are symmetrically and evenly spaced. Furthermore,
we can choose these values such that the three rail edges are of length unity and
so that the one-hop lengths are also all of length unity, and the two-hop lengths
are slightly longer. We call such a tetrahelix a two-class tetrahelix.

@@

Now that we have shown that any optimal tetrahelix vertices
are on helices of the same axes and pitch, we see that the vertices 
of any optimal tetrahelix will lie on a cylinder, or a circle when the axis dimension
is projected out. Therefore it is reasonable to now speak of the singular \emph{radius} $r$
of a tetrahelix as the radius of the cylinder.
We can now go on to
the harder proof about where vertices occur along the $z$-axis.

@@

    Any attempt to perturb any rail in either $z$ direction lengthens one two-hop edge to $t'$, where $t' > t$
    and shortens one one-hop edge $o' < o$. Let $B = \{o',t' \} \cup A$ be the edge lengths of such a
    perturbed tetrahelix.
    The minimax of $B$ is greater than the minimax of $A$ since there is a single rail length which cannot be both greater
    than $t'$ and $o'$ and less than $t'$ and $o'$.
    Therefore, any optimal tetrahelix has all one-hop edges between all rails equal to each other,
    all two-hop edges equal to each other, the $z$ distances between rails equal, and therefore
    $d/3$ from each other.
\end{proof}

@@

\begin{align*}
  1^2 &= d^2 + (2 r \sin{ \rho / 2})^2 \\
%%  1 &= d^2 + 4 r^2 (\sin{ \rho / 2})^2 \\
  d^2 &= 1 - 4 r^2 (\sin{ \rho / 2})^2    \numberthis  \label{railangle} \\
\end{align*}

@@

Rewriting our formulation in terms of $\rho$:
\begin{align*}
\numberthis  \label{eq:general}  
\vec{H}_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) &= 
 \left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ d_{\rho} (n + c /3) $
  \end{tabular}
  \right ] \\
\text{where: } &
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
    $\chi \in \{-1,1\}$ \\  
\end{tabular}
\end{align*}

@@

\begin{equation*}
  j_{\rho} = \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)
\end{equation*}
\begin{equation*}
  k_{\rho} = (\sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3}))^2)
\end{equation*}

@@

By graph inspection using Mathematica (\url{https://github.com/PubInv/tetrahelix/blob/master/tetrahelix.nb}), we see the partial derivative of this with respect to
radius $r$ is always negative, for any $\rho \leq \rho_{bc}$. (When the rail angle approaches
$\pi$, corresponding to going almost to the other side of the tetrahelix, this is not necessarily true, hence the
limitation in our statement of the theorem is meaningful.)
Since the partial derivative of $\text{two-hop}(r) / \text{one-hop}(r)$ with respect to the
radius $r$ is negative for all $\rho$ up until $\rho_{bc}$, this ratio goes down
as the radius goes up, and 
we minimize the maximum edge-length ratio by choosing the largest radius
up until one-hop $= 1$, the rail-edge length. If we attempted to increase the radius further
we would not be optimal, because the ratio $\frac{\text{two-hop}(r)}{1}$ would become the
largest ratio in our set of ratios \cref{eq:ratioset}.

@@

\begin{align*}
  1 &=  \sqrt{\frac{1}{9}  + r_{opt}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  r_{opt} &= \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }} \numberthis  \label{eqrhoopt} \\
\end{align*}

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 (\frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }})^2 (\sin{ \rho / 2})^2   \\
%%  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
%%  d_{opt}^2 &= 1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}

@@

Since the axes are parallel, we may define the \emph{inradius}, represented by the letter $i$, of a
tetrahelix to be the radius of the largest
cylinder parallel to this axis that is surrounded by each tetrahelix and pentrated by no edge.

@@

In this way a rail angle can be chosen for any desired (sufficiently large) pitch, yielding
the optimum radius, the one-hop length, and the two-hop length an engineer needs to
construct a physical structure.

@@

Perhaps surprisingly, the optimal untwisting is accomplished only by
changing the length of the two-hop member, leaving the one-hop member
and rail length equivalent within this continuum.\footnote{Before
deriving Equation \eqref{eqrhoopt}, we created a continuum by
using a linear interpolation between the optimal radius for the
Equitetrabeam and the BC Helix. This minimax optimum of this simpler
approach was at most 1\% worse than the optimum computed by
\eqref{eqrhoopt}.}
 However, it should
be noted that an engineer or architect may also use $\vec{H}_{general}$ 
directly and interactively  \url{https://pubinv.github.io/tetrahelix/},
and that minimax length optimality is a
mathematical starting point rather than the final word on the beauty and utility of
physical structures. For example, a structural engineer might increase
radius past optimality in order to resist buckling.
%% Add a figure here of a deep beam and of an arch.

>>>>

\begin{abstract}
  The Boerdijk-Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra forming a helical column.  Considering the edges of
  these tetrahedra as structural members, the resulting structure is attractive and
  inherently rigid, and therefore interesting to architects,
  mechanical engineers, and roboticists.  A formula is developed that matches the
  visually apparent helices forming the outer rails of the BC helix.
  This formula is generalized to a formula convenient to designers.
  Formulae for 
  computing the
  parameters that give edge-length minimax-optimal tetrahelices
  are given, defining a continuum of optimum tetrahelices of varying curvature.
  The endpoints of this continuum are the BC helix and
  a structure of zero curvature, the \emph{equitetrabeam}.
  Numerically finding the rail angle from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed. 
  An interactive tool for such design and experimentation is provided: \url{https://pubinv.github.io/tetrahelix/}.
  A formula for the inradius of optimal tetrahelices is given.
  Utility for static and variable geometry
  truss/space frame design and robotics is discussed.
\end{abstract}

@@

Imagining \cref{fig:closeup,fig:closeuporthog} as a static mechanical structure,
we observe that it is useful to the mechanical engineer or
roboticist because the structure is an inherently rigid,
omni-triangulated space frame, which is
mechanically strong.
Then we can imagine that each static edge is replaced with an
actuator that can dynamically become shorter or longer in response to electronic control,
and the vertices are joints that support sufficient angular displacement
for this to be possible. An example of such a machine is a Tetrobot,
shown in \cref{fig:tetrobot}.

A BC helix does not rest stably on a plane. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
developed here is available in JavaScript and demonstrated by an interactive
design website \url{https://pubinv.github.io/tetrahelix/}\cite{readtetrahelix},
from which \cref{fig:closeup,fig:closeuporthog,fig:series,fig:naming,fig:equitetrabeam}
are taken.

@@

The apparent rotation of a vertex on an outer-edge, that is $\vec{V}(n)$ relative from $\vec{V}(n+3)$
for any integer $n$
in \eqref{graycoxeter}, is $3 \theta_{bc} - 2\pi$.

This formula defines a helix, but it is not any of the apparent helices, or \emph{rail} helices, of the
BC helix, but rather one that winds much more rapidly through all
nodes. To a designer of tetrahelices, it is more natural to think of
the three helices which are visually apparent, that is, those three
which are closely approximated by the outer edges or rails of
the BC helix. We think of each of these three rails as being a different color: red, blue, or yellow.
This situation is illustrated in \cref{fig:helixcomparison}, wherein the black helix represents that
generated by \cref{graycoxeter},
and the colored helices are generated by \cref{eq:colored}.

@@

Such a helix can be written:
\begin{equation}
  \label{eq:colored}
 \vec{H}_{BCcolored}(n,c) =
\left [
  \begin{tabular}{c}
   $ r_{bc}  \cos{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})}  $\\
   $ r_{bc} \sin{((3 \theta_{bc} - 2 \pi)n + c  \theta_{bc})} $\\
   $ 3 h_{bc} (n + c/3)  $
  \end{tabular}
  \right ],
\text{where}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta_{bc} = \arccos(-2/3) $ \text{ .}\\
  \end{tabular}      
\end{equation}

@@

The $\vec{H}_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho_{bc}$ rather than $\theta_{bc}$.
Intuitively we seek an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$.
We expand 
the expressions $\theta_{bc}$ and $\rho_{bc}$ in \eqref{eq:colored} and seek to isolate the term $c 2\pi/3 $. Thus:
\begin{align*}
 c \theta_{bc}  &=   \text{\{we aim for 3 in denominator, so we split...\}} \\
    (c/3)  (3 \theta_{bc})  &=   \text{\{we want $2\pi$ in numerator, so add canceling terms...\}} \\
 (c/ 3) ((3 \theta_{bc} - 2 \pi)  + 2 \pi) &= \text{\{definition of $\rho_{bc}$\ ...\}} \\  
  (c / 3) \rho_{bc}  + c 2 \pi /3 &=  \text{\{algebra...\}} \\  
c  ( \rho_{bc} +  2 \pi) /3 \text{ .} \\
\end{align*}

@@

The pitch is less than the number of tetrahedra because the tetrahedra
edges are not parallel to the axis of the tetrahelix.  It is a famous and interesting result
that the pitch is irrational. A BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of developing a rational period by arranging solid tetrahedra by relaxing the face-to-face matching
has been explored\cite{sadler2013periodic}. 
We develop below slightly irregular edge lengths that support, for example, a pitch of precisely 12
tetrahedra in one revolution which would allow an architect to
design a pleasing column having the top and bottom tetrahedra in the same relationship to the capital and the basis to the
viewer.

@@

In the proof below we find useful to consider projection diagrams that are the axial projection
of a tetrahelix onto the $XY$-plane. \Cref{fig:projectiondiagram} is an example of such a diagram.
\begin{lemma}
  If the rail angle $0 < \rho < \pi$ is a rational multiple of $\pi$, then the projection of
  edges in a helix of that rail angle  along the $z$-axis onto the $XY$-plane form a
  regular polygon of 3 or more sides, or else they fill in a complete circle.
  \label{lemma:ngon}
\end{lemma}
\begin{proof}
  All points lying on a helix projected along the axis lie on a circle in the $XY$-plane.
  Helices are periodic in the $z$ dimension modulo $2\pi$.
  If $2\pi/\rho$ is irrational, the projection onto the $XY$-plane will contain
  an unbounded number of points on a circle.
  If and only if $2\pi/\rho$ is rational, the projection onto
  the $XY$-plane will contain a finite number of points.
  Because $\pi$ is transcendental and irrational,
  $2\pi/\rho$ is rational if and only $\rho = a\pi/b$, where $a$ and $b$ are integers
  and without loss of generality $a$ and $b$ are coprime.
  Since $\rho < \pi$, therefore $a < b$. Also, $\rho > 0$, therefore $a > 0$.
  The number of points in the projection is $2b$ if $a$ is odd, and $b$ if $a$ is even.
  This polygon has at least 3 sides, since either $\rho$ is irrational or $b > a$, and therefore
  $b >= 2$. If $a/b = 1/2$, the projection is a square, which has four sides. 
\end{proof}

\begin{theorem}
  Any optimal tetrahelix with a rail angle of magnitude less than $\pi$ has all three
  axes conincident.
  \label{thm:conincident}
\end{theorem}
\begin{proof}
  Case 1: Suppose that $\rho$ is zero.
  Each helix has zero curvature, that is, it is a straight line. These lines are eqivalent
  to some three degenerate helices with conincident axes, possibly with different radii, so long as there is a phase
  term in the defintion of the helix, as in \eqref{eq:colored}. We later show the
  radii must be eqivalent.

  Case 2: Suppose that $\rho$ is positive but less than $\pi$.
  In this case each rail helix has
  curvature. The projection of points in the $XY$ plane creates a figure
  guaranteed to have points on either side of any line through the axis of such
  a helix, because the figure is either an $n$-gon or a circle by \cref{lemma:ngon}.
  We show that the three helices share a common axis.

  Without loss of generality define the Red helix to have its axis on the $z$-axis.
  Since there must be at least one Red-to-Yellow or a Red-to-Blue edge that is either a minimum or a maximum,
  without loss of generatlity define the Blue helix to be a helix that
  has an edge connection to the Red helix that is either a maximum or a minimum.
  Let $B'$ be a translation in the $XY$-plane of the blue helix $B$  so that its axis is the $z$-axis and
  conincident with the red helix $R$. Let $D$ be the distance between the axis of the Blue helix
  $B$ and $B'$. We will show that if $D>0$ then $B$ ``wobbles'' in a way that cannot be optimal.
  Define a wobble vector by:
  \begin{equation*}
    \vec{W}(n) = \vec{B}(n) - \vec{B}'(n) \text{ .}
  \end{equation*}
  where $\vec{B}(n)$ and $\vec{B'}(n)$ is the vector
  $\begin{bmatrix}
    x \\
    y 
  \end{bmatrix}$
  for the projection of the $n$th vertex of $B$ and $B'$. 
  Note that $|| \vec{R}(n) - \vec{B'}(n+k)||$ (the Euclidean distance of the vertices)
  is a constant for any $k$, because $R$ and $B'$ have the
  same pitch and the same axis, even if they do not have the same radius.

@@

    This process can be carried out on both the Blue and Yellow helices
    (perhaps simultaneously) until $\vec{W}(n)$ is
    zero for both, finding a tetrahelix of improved overall minimax value at each step.
    So a tetrahelix is optimal only when $\vec{W}(n) = 0$, and therefore when $D=0$ and
    $\vec{B}(n) = \vec{B'}(n)$, and all three axes are coincident.
\end{proof}

@@

In principle any three helices generated with \cref{eq:triple}
has at most nine distinct edge length classes. Each edge that connects
two rails potentially has a longer length and shorter length we denote with
a $+$ or $-$. So the classes are $\{ RR, BB, YY, RB_+, RB_-, BY_+, BY_-, RY_+, RY_- \}$.
If when projecting all vertices onto the $z$-axis (dropping the $x$ and $y$ coordinates), the interval
defined by the $z$ axis value of its endpoints contains no other vertices,
we call it a \emph{one-hop} edge, and if it does contain another vertex we
call it a \emph{two-hop} edge, as illustrated in \cref{fig:naming}.
Then there are 
3 rail edges $\{ RR, BB, YY\}$,
3 one-hop lengths $\{ RB_-, BY_-, RY_- \}$ between each pair of 3 rails,
and 3 two-hop
lengths $\{ RB_{+}, BY_+, RY_+ \}$ between each pair of 3 rails,
where the two-hop length is at least the one-hop length.
However, if we generate the three helices symmetrically with \cref{eq:triple},
many of these lengths will be the same.
In fact, it is possible that there
will be only two distinct such classes, or even one, in the purely regular BC helix.

\begin{theorem}
  Optimal tetrahelices have the same radii for all three helices.
\end{theorem}
\begin{proof}
To prove this we exhibit a symmetric tetrahelix (not yet shown to be optimal) which
happens to be a triple helix, that has
the property that all rail edges are equal to all one-hop edges and all two-hop
edges are equal to each other. 
Observe that although we have not get given the formula for the
radii of such a triple helix, we observe there are some values for $r$ and $\alpha$, and $\phi$
in \cref{eq:triple}
for which all the three helices are symmetrically and evenly spaced. Furthermore,
we can choose these values such that the three rail edges are of length $1$ and
so that the one-hop lengths are also all of length $1$, and the two-hop lengths
are slightly longer.

@@

Now that we have shown that on any optimal tetrahelix the vertices
are on helices of the same axes and pitch, we see that the vertices 
of any optimal tetrahelix will lie on a cylinder, or a circle when the axis dimension
is projected out. Therefore it is reasonable to now speak of the singular \emph{radius} $r$
of a tetrahelix as the radius of the cylinder.
We can now go on to
the harder proof about where vertices occur along the $z$-axis.

@@

    Any attempt to perturb any rail in either $z$ direction lengthens one two-hop edge to $t'$, where $t' > t$
    and shortens one one-hop edge $o' < o$. Let $B = \{o',t' \} \cup A$ be the edge lengths of such a
    perturbed tetrahelix.
    The minimax of $B$ is greater than the minimax of $A$ since there is a single rail length which cannot be both greater
    than $t'$ and $o'$ and less than $t'$ and $o'$.
    Therefore, any optimal tetrahelix has all one-hop edges between all rails equal to each other, and
    all two-hop edges equal to each other, the $z$ distances between rails equal. Therefore vertices are
    $d/3$ from each other on the $z$-axis.
\end{proof}

@@

\begin{align*}
  1^2 &= d^2 + (2 r \sin{ \rho / 2})^2 , \text{ or }\\
%%  1 &= d^2 + 4 r^2 (\sin{ \rho / 2})^2 \\
  d^2 &= 1 - 4 r^2 (\sin{ \rho / 2})^2    \numberthis  \label{railangle} \\
\end{align*}

@@

Rewriting our formulation in terms of $\rho$:
\begin{align*}
\numberthis  \label{eq:general}  
\vec{H}_{general}(\chi,n,c,\rho,d_{\rho},r_{\rho}) &= 
 \left [
  \begin{tabular}{c}
   $ r_{\rho} \cos(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ r_{\rho}  \sin(\chi \cdot (n \rho + c(\rho +  2 \pi) /3)) $\\
   $ d_{\rho} (n + c /3) $
  \end{tabular}
  \right ] \\
\text{where: } &
\begin{tabular}{c}
  $   1 = d_{\rho}^2 + 4 r_{\rho}^2 (\sin{ \rho / 2})^2 $ \\
    $\chi \in \{-1,1\}$ \text{ .} \\  
\end{tabular}
\end{align*}

@@

\begin{equation*}
  j_{\rho} = \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2
\end{equation*}
\begin{equation*}
  k_{\rho} = \sin^2(2\rho/3 + \frac{4\pi}{3})  + (1 - \cos(2\rho/3 + \frac{4\pi}{3}))^2
\end{equation*}

@@

By graph inspection using Mathematica (\url{https://github.com/PubInv/tetrahelix/blob/master/tetrahelix.nb}), we see the partial derivative of this with respect to
radius $r$ is always negative, for any $\rho \leq \rho_{bc}$. (When the rail angle approaches
$\pi$, corresponding to going almost to the other side of the tetrahelix, this is not necessarily true, hence the
limitation in our statement of the theorem is meaningful.)
Since the partial derivative of  $\frac{\text{two-hop}(r)}{\text{one-hop}(r)}$ with respect to the
radius $r$ is negative for all $\rho$ up until $\rho_{bc}$, this ratio goes down
as the radius goes up, and 
we minimize the maximum edge-length ratio by choosing the largest radius
up until one-hop $= 1$, the rail-edge length. If we attempted to increase the radius further
we would not be optimal, because the ratio $\frac{\text{two-hop}(r)}{1}$ would become the
largest ratio in our set of ratios \cref{eq:ratioset}.

@@

\begin{align*}
  1 &=  \sqrt{\frac{1}{9}  + r_{opt}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  r_{opt} &= \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }} \text{ .} \numberthis  \label{eqrhoopt} \\
\end{align*}

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 (\frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }})^2 (\sin{ \rho / 2})^2   \\
%%  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
%%  d_{opt}^2 &= 1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }}  \text{ .}  \numberthis  \label{dopt}  \\      
\end{align*}

@@

Since the axes are parallel, we may define the \emph{inradius}, represented by the letter $i$, of a
tetrahelix to be the radius of the largest
cylinder parallel to this axis that is surrounded by each tetrahelix and penetrated by no edge.

@@

In this way a rail angle can be chosen for any desired (sufficiently large) pitch, yielding
the optimum radius, the one-hop length, and the two-hop length that an engineer needs to
construct a physical structure.

@@

Perhaps surprisingly, the optimal untwisting is accomplished only by
changing the length of the two-hop member, leaving the one-hop member
and rail length equivalent within this continuum.\footnote{Before
deriving Equation \eqref{eqrhoopt}, we created a continuum by
using a linear interpolation between the optimal radius for the
Equitetrabeam and the BC Helix. This minimax optimum of this simpler
approach was at most 1\% worse than the optimum computed by
\eqref{eqrhoopt}.}
 However, it should
be noted that an engineer or architect may also use $\vec{H}_{general}$ 
directly and interactively via \url{https://pubinv.github.io/tetrahelix/},
and that minimax length optimality is a
mathematical starting point rather than the final word on the beauty and utility of
physical structures. For example, a structural engineer might increase
radius past optimality in order to resist buckling.
%% Add a figure here of a deep beam and of an arch.

If an equitetrabeam were actually
used as a beam, an engineer might start with the optimal tetrabeam and
dilate it in one dimension to stiffen the beam by deepening it. Similarly, simple
length changes curve the equitetrabeam into an arch.
The ``colored'' approach of \cref{eq:general} exposes these possibilities
more than the approach of \cref{graycoxeter}.
